---
title: "GDP"
author: "Guibril Ramde"
format: html
editor: visual
---

## Approach

In this project, I selected the **Worldwide GDP dataset** from Discussion 5A. The dataset is not initially structured in a tidy format because the GDP values for different years are stored across multiple columns rather than in a single variable column. Therefore, the first step of this analysis will be to clean and restructure the dataset using tidy data principles.

Using tools from the **tidyverse** package, I will transform the dataset into a tidy format where each variable has its own column, each observation corresponds to a single row, and each type of observational unit forms a table. This process will involve reshaping the data so that the year values become a single column and the GDP values become another column.

After the dataset has been tidied, I will perform an exploratory analysis to examine economic patterns across countries. In particular, I will analyze how GDP trends evolve over time for different countries and investigate possible relationships with factors such as population growth and birth rates. I will also explore how these economic indicators have changed over the past decade and identify notable global or regional patterns.

Finally, I will use visualization tools such as **ggplot2** to present the results and highlight key insights from the data.

```{r}
library(dplyr)
library(tidyr)
library(readr)
library(tidyverse)

load_data <- "https://raw.githubusercontent.com/japhet125/Project2-Data-Science/refs/heads/main/GDP.csv"
get_all_data <- read_csv(load_data)
get_all_data
```

We will use ggplot to visualize and analize the dataset

**3.2 Data Import and Tidying**

```{r}
clean_data <- get_all_data %>%
  pivot_longer(
    cols = `1960`:`2022`,
    names_to = "year",
    values_to = "gdp_Amount"
  )
clean_data
```

**Normalizing variable structure and Rename variable**

```{r}
clean_data %>%
  rename(
    Country_Code = `Country Code`
  )
```

Handling missing data, since they do not provide measurable data or any informations

```{r}
clean_data %>%
  filter(!is.na(gdp_Amount))
```
